Embracing the Laws of Physics: \\ Three Reversible Models of Computation

### 2018-07-05 survey/survey.tex
add an new section on internal language in section 4. Draw the monster commutative diagram. Second one coming later.

%%%%%%%%%
\subsection{Example}

@@

% \jc{The above is, up to uses of assoc, a linear notation for a
% commutative diagram. So draw it.}

>>>>

%%%%%%%%%
\subsection{Example}\label{sec:level2-example}

@@

\subsection{Internal Language}

Recalling that the $\lambda$-calculus arises as the internal language
of Cartesian Closed Categories (Elliott~\cite{Elliott-2017} gives a particularly
readable account of this), we can think of $\Pi$ in similar terms, but
for symmetric Rig Groupoids instead. For example, we can ask what does
the derivation in Sec.~\ref{sec:level2-example} represent? It is
actually a ``linear'' representation of a 2-categorial commutative
diagram! In fact, it is a painfully verbose version thereof, as it
includes many \emph{refocusing} steps because our language does not
build associativity into its syntax. Categorical diagrams usually do.
Thus if we rewrite the example in diagrammatic form, eliding all uses
of associativity, but keeping explicit uses of identity transformations,
we get that \AgdaFunction{swap{-}fl2⇔swap{-}fl1} represents

\newcommand{\idd}{\mathit{id}\leftrightarrow}
\newcommand{\idf}{\mathit{id}\Leftrightarrow}

\begin{tikzcd}[column sep=normal, row sep=small]
 && (a+c)+b \arrow [r, "\swapp \oplus\idd", ""{name=U, below}] & (c+a)+b \arrow [dr, "\assocrp"] && \\
 & a+(c+b) \arrow [ur, "\assoclp"] & & & c+(a+b) \arrow [dr, "\idd\oplus\swapp"] &  \\
a+(b+c) \arrow [ur, "\idd\oplus\swapp"] \arrow [rr, "\assoclp\odot\swapp"] 
  \arrow [ddrr, "\assoclp\odot\swapp"]
  \arrow [dddrr, swap, "\assoclp\odot\swapp"]
    &&
    c+(a+b) \arrow [r, swap, "\assoclp", ""{name=D, above}] 
    & |[alias=Z]| (c+a)+b \arrow [r, "\assocrp"] &c+(a+b) \arrow [r, "\idd\oplus\swapp"] & c+(b+a) \\
 &&&&& \\
 & & c+(a+b) \arrow [rr, swap, "\idd", ""{name=DD, above}]
             \arrow [d, Rightarrow, "\idf\, \mathit{idl}\odot{l}"] && 
    c+(a+b) \arrow [ruu, swap, "\idd\oplus\swapp"] & \\
 && c+(a+b) \arrow [rrruuu, bend right = 30, swap, "\idd\oplus\swapp"] && \\
 \arrow[Rightarrow, from=U, to=D, "\mathit{hexagon}\oplus{r}\, \boxdot\, \idf"]
 \arrow[Rightarrow, from=Z, to=DD, "\idf\boxdot\mathit{linv}\odot{l}\,\boxdot\,\idf"]
\end{tikzcd}

\noindent For some, the above diagram will be clearer --- it is only three layers
high rather than nine! Others will prefer the more programmatic feel of the
original definition.

We would be remiss in letting the reader believe that the above is ``the''
categorical diagram that would be found in categorical textbooks. Rather,
congruence would be used to elide the $\idf$. The various arrows would also
be named differently --- our \assoclp is often named $\alpha$,
\assocrp is $\alpha^{-1}$, $\swapp$ is $B$ (always with subscripts).
